### 2016-08-05 capability-machine.tex
.

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2) \in \phi_1
\]
\begin{definition}[Heap satisfaction/erasure]
\[
\heapSat[\hs]{n}{W} \text{ iff }
\begin{aligned}
&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp \\
&\quad\begin{gathered}
   hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}\\
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{gathered}
\end{aligned}
\]
\end{definition}

A standard heap invariant that insures all values in the region are in the value  relation:
\[
  \iota_{\var{start},\var{end}} \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{base},\var{end}}(W) & \defeq (s,=,H_{\var{base},\var{end}}) \\
  H_{\var{base},\var{end}}\; s\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
                             & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
\end{align*}
Another standard heap invariant that ensures all the words in the region are 0.
\[
  \iota_{\var{start},\var{end}}^0 \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq (s,=,H_{\var{base},\var{end}}^0) \\
  H_{\var{base},\var{end}}^0\; s\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
\end{align*}

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionNames \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nsubeq[n-1] \iota_{\start',\addrend'} \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
\end{align*}

@@

\begin{lemma}[Heap satisfaction uniformity]
  \begin{align*}
    & \forall n,n',\heap, W \ldotp\\
    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
  \end{align*}
\end{lemma}

@@

@@

@@

@@

>>>>

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2) \in \phi_1
\]

n-subset for regions
\[
(s_1,\phi_1,H_1) \nsubeq (s_2,\phi_2,H_2) \text{ iff } 
\begin{gathered}
(s_1,\phi_1) = (s_2,\phi_2)\\
 \text{ and }\\
 \forall \hat{W} \ldotp H_1 \; s_1 \; \hat{W} \nsubeq H_2 \; s_2 \; \hat{W}
\end{gathered}
\]
\begin{definition}[Heap satisfaction/erasure]
\[
\heapSat[\hs]{n}{W} \text{ iff }
\begin{aligned}
&\exists R : \dom(W) \rightarrow \HeapSegments \ldotp \\
&\quad\begin{gathered}
   hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}\\
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{gathered}
\end{aligned}
\]
\end{definition}

A standard heap invariant that insures all values in the region are in the value  relation:
\[
  \iota_{\var{start},\var{end}} \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{base},\var{end}}(W) & \defeq ((\var{base},\var{end}),=,H_{\var{std}}) \\
  H_{\var{std}}\; (\var{base},\var{end})\; W         & \defeq \{\npair{\hs} \mid \forall \addr \in [\var{base},\var{end}] \ldotp \hs(\addr) \in \stdvr(\roll \; W) \\
                             & \qquad \qquad \qquad \land \dom(\hs) = [\var{base},\var{end}]\}
\end{align*}
where $s$ is some state. It does not matter which one because the predicate does not use. 
% This is unlike the \lambda_JS paper where the state in the standard island is a location, namely the location that the standard island governs.

Another standard heap invariant that ensures all the words in the region are 0.
\[
  \iota_{\var{start},\var{end}}^0 \; : \; \Regions
\]
\begin{align*}
  \iota_{\var{start},\var{end}}^0\; s\; W & \defeq ((\var{base},\var{end}),=,H_{\var{std}}^0) \\
  H_{\var{std}}^0\; (\var{base},\var{end})\; W     & \defeq\{ \npair{\heapseg} \mid \dom(\heapseg) = [\var{start},\var{end}] \land \\
                                                                & \qquad \forall \addr \in \dom(\heapseg) \ldotp \heapseg(\addr) = 0  \}
\end{align*}

\begin{align*}
  \readCond{n,W,\start,\addrend} =        & \;\exists r \in \RegionNames \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp\\
                                          & \;\qquad W(r)\nsubeq[n-1] \iota_{\start',\addrend'}(r) \\ & \\
  \writeCond{n,W,\start,\addrend} =       & \; \exists r \in \RegionNames \ldotp \\
                                          & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp W(r) \nequal[n-1] \iota_{\start',\addrend'} \\ & \\
  \execCond{n,W,\start,\addrend,\perm} =  & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \forall \addr \in [\start,\addrend] \ldotp \\
                                          & \; \qquad \npair[n']{(\perm,\start,\addrend,\addr)} \in \stder(W') \\ & \\
  \entryCond{n,W,\start,\addrend,\addr} = & \; \forall n' < n \ldotp \forall W' \future W \ldotp \\
                                          & \; \quad \npair[n']{(\exec,\start,\addrend,\addr)} \in \stder(W') \land \\
\end{align*}

@@

\subsection{Well-foundedness lemmas}
\begin{lemma}[Heap satisfaction uniformity]
  \begin{align*}
    & \forall n,n',\heap, W \ldotp\\
    & \quad \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
  \end{align*}
\end{lemma}

@@

\begin{lemma}[Read condition monotone in world]
  \begin{align*}
    &\forall n,W,W',\start, \addrend \ldotp \\
    &\quad  \readCond{n,W,\start,\addrend} \land \\
    &\quad  W' \future W \\
    &\qquad \Rightarrow \readCond{n,W',\start,\addrend}
  \end{align*}
\end{lemma}
\begin{proof}
Let $n$,$W$, $W'$, $b$, and $e$ be given where $W' \future W$. Assume $\readCond{n,W,\start,\addrend}$, this give a $r$ and $[b',e'] \supseteq [b,e]$ such that $W(r) \nsubeq \iota_{b',e'} (r) = ((b',e'),=,H_{\var{std}})$. This gives
\begin{align*}
  &W(r).\phi = = \\
  &W(r).s    = (b',e') \\
  &\forall \hat{W} \ldotp W(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
\end{align*}
$W' \future W$ gives
\begin{align*}
  &W'(r).\phi = W(r).phi \\
  &(W'(r).s,W(r).s) \in W(r).\phi \\
  &W'(r).H = W(r).H \\
 \end{align*}
Combining these two sets of properties line for line, we get the following for $W'$:
\begin{align*}
  & W'(r).\phi = = \\
  & W'(r) = (b',e') \\
  &\forall \hat{W} \ldotp W'(r).H\; (b',e') \; \hat{W} \nsubeq[n-1] H_{\var{std}}\; (b',e') \; \hat{W}
\end{align*}
These are exactly the properties neede to get $W'(r) \nsubeq \iota_{b',e'} (r)$ which is exactly the result we want.
\end{proof}

@@

\begin{lemma}[Read-write condition monotone in world]
  \begin{align*}
    & \forall n,W, W',\start, \addrend \ldotp\\
    &\quad \writeCond{n,W,\start,\addrend} \land\\
    &\quad  W' \leq W\\
    &\qquad \Rightarrow \writeCond{n,W',\start,\addrend}
  \end{align*}
\end{lemma}
\begin{proof}
  Similar to the proof of the monotonicity of the read condition.
\end{proof}

@@

\begin{lemma}[Execute condition uniformity]
  \begin{align*}
    &\forall n,W, W',\start, \addrend, \perm \ldotp \\
    &\quad  \execCond{n,W,\start,\addrend,\perm} \land \\
    &\quad  W' \leq W \\
    &\qquad \Rightarrow \execCond{n,W',\start,\addrend,\perm}
  \end{align*}
\end{lemma}
\begin{proof}
  Follows directly from definition.
\end{proof}

@@

\begin{lemma}[Entry condition uniformity]
  \begin{align*}
    & \forall n,W,W',\start, \addrend, \addr \ldotp\\
    &\quad  \entryCond{n,W,\start,\addrend,\addr} \land \\
    &\quad  W' \leq W \\
    &\qquad \Rightarrow \entryCond{n,W',\start,\addrend,\addr}
  \end{align*}
\end{lemma}
\begin{proof}
  Follows directly from definition
\end{proof}
